---
 title: "Results on census-eligible"
 author: "Andrea Brizzi"
 date: "`r Sys.Date()`"
 output: 
    html_document:
        theme: journal
        highlight: zenburn
        toc: true
        toc_depth: 2
        toc_float: true
 
---

```{r, echo=FALSE, message=FALSE, results='hide'}

library(data.table)
library(ggplot2)
library(knitr)

gitdir <- here::here()
source(file.path(gitdir, "R/paths.R"))

# helpers
classify_type_by_name <- function(label){

    out <- fcase(
        label %like% 'hivprev|prevl', 'hivprevalence',
        label %like% 'suppofhiv', 'suppressionamonghiv',
        label %like% 'suppofpop', 'viraemiainpop',
        default = NA_character_
    )

}

get_plotname <- function(label){
    gsub(
        pattern="round1[0-9]|hivprev|prevl|suppofhiv|suppofpop|\\_|\\.png$|\\.pdf$",
        replacement = "",
        label)
}

insert_image <- function(path, dir=out.dir.figures){
    full.path <- file.path(dir, path)
    sprintf("\n![](%s)\n", full.path)
}

# dictionary to map plot labels to proper section names 
type2text <- function(x){

    dict <- c(
        'logratioftpvsnnonbyroundcomm' = "Posterior log ratio of participants vs first-time participants", 
        'barenumviraemiccensel' = "Expected number of individuals exhibiting viraemia among census-eligible",
        'lineecontrviraemiccensel' = "Expected prevalence of viraemia among census-eligible",
        'contribagegenderbyroundcomm' = "Gender-age contributions by round and community",
        'contribagegroupgenderbyroundcomm' = "Gender-agegroup contributions by round and community",
        'fitbyftpstatus' = 'Posterior fits by firt-time participation status',
        'fitbyroundcommgender' = 'Aggregated posterior fits by round, community type and gender',
        'hivprevalence' = 'HIV prevalence',
        'suppressionamonghiv' = 'Suppression in HIV',
        'viraemiainpop' = 'Prevalence of viraemia',
        NULL
    )
    dict[x]
}
```

First let us load the figures obtained in the analysis: 

```{r load-data}
opts_vec <- c(
    "viremic.viral.load", "detectable.viral.load", "out.dir.prefix", 
    "indir", "round", "jobname", "only.firstparticipants")
args <- args[names(args) %in% opts_vec]

VL_DETECTABLE <- args$vl.detectable
VIREMIC_VIRAL_LOAD <- args$viremic.viral.load

# output directories
out.dir <- args$out.dir.prefix
out.dir <- file.path(out.dir, paste0("vl_", VIREMIC_VIRAL_LOAD, "_joint"))
out.dir.figures <- file.path(out.dir, "figures")
out.dir.tables <- file.path(out.dir, "tables")

paths <- list.files(out.dir.figures, pattern='png$')
dplots <- data.table(
    PATH = paths, 
    PLOT = get_plotname(paths),
    ROUND = gsub('^.*round(1[0-9]).*$', '\\1', paths) |> as.integer() |> suppressWarnings(), 
    TYPE = classify_type_by_name(paths)
)

paths <- list.files( file.path(args$out.dir.prefix, 'figures'), pattern='png$')
dplots_data <- data.table(
    PATH = paths,
    PLOT = get_plotname(paths),
    ROUND = NA_character_,
    TYPE = classify_type_by_name(paths)
)
```

The below shows the postprocessing plots, for analyses on three different measures of interest:

- prevalence of HIV.
- viral suppression among HIV infected individuals.
- prevalence of viraemia.

We have data for participants to the Rakai Community Cohort Study, however we wanted to generalise our findings to the entire population of those who were census-eligible (aged 15-50, no intention to migrate).
To model non-participants, we assumed that **first-time-participants are an homogeneous subset of non participants**.
As such we: 

1. Fitted the three models twice: on all RCCS participants, and then exclusively including first-time-participants (FTP)
2. Aggregated the two model fits using participation rates as weights.

Note that FTP is included in both analyses, as FTP are a subset of the RCCS participants.

```{r plot-order, echo=FALSE, hidden=TRUE, include=FALSE}
ordered_labels <- c( 
    'fitbyftpstatus',
    'logratioftpvsnnonbyroundcomm',
    'fitbyroundcommgender',
    'contribagegenderbyroundcomm' ,
    'contribagegroupgenderbyroundcomm',
    'barenumviraemiccensel',
    'lineecontrviraemiccensel'
)
dplots[, PLOT := ordered(PLOT, levels=ordered_labels)]
setkey(dplots, PLOT, ROUND)
```


```{r pyramid-plots, echo=FALSE, hidden=TRUE, message=FALSE, results='asis', fig.height=12, fig.width=14}

sprintf("\n## %s {.tabset} \n", "Raw data: population pyramids") |> cat()
sprintf("\nLet us start by plotting the raw data in form of population pyramids.\n") |> cat()

# set order 
dplots_data[PLOT %like% 'pyramid', 
    ORDER := fcase(
        PLOT %like% 'NeligibleN', 1,
        PLOT %like% 'NparticipantsN', 2,
        PLOT %like% 'NhivpN', 3
) ]
setkey(dplots_data, ORDER)

dplots_data[PLOT %like% 'pyramid', {
    toprint <- gsub("pyramidN(.*)N(.*)$","\\1 : \\2",PLOT)
    sprintf("\n### %s \n", toprint) |> cat()
    insert_image(PATH[1], dir=file.path(args$out.dir.prefix, 'figures')) |> cat()
}, by='PLOT']

```

```{r show-plots, echo=FALSE, hidden=TRUE, message=FALSE, results='asis', fig.height=12, fig.width=14}
{
    dplots[, {
        
        # get title of section
        ascPLOT = unique(as.character(PLOT))
        sprintf("\n## %s {.tabset } \n", unique(type2text(ascPLOT)) ) |> cat()

        uniq_type <- unique(TYPE)
        uniq_round <- unique(ROUND)
        many_type <- length(uniq_type) > 1
        many_round <- length(uniq_round) > 1

        if ( many_type & many_round ){

            for (type in uniq_type){
                sprintf("\n### %s {.tabset} \n", type2text(type)) |> cat()

                for (round in uniq_round){
                    sprintf("\n#### Round %s \n", round) |> cat()

                    idx <- which(ascPLOT == ascPLOT & ROUND == round & TYPE == type)
                    insert_image(PATH[idx]) |> cat()
                }
            }
        }else if(many_type){
            for (type in unique(TYPE)){
                idx <- which(ascPLOT == ascPLOT & TYPE == type)
                sprintf("\n### %s \n", type2text(type)) |> cat()
                insert_image(PATH[idx]) |> cat()
            }
        }else if(many_round){
            for (round in unique(ROUND)){
                idx <- which(ascPLOT == ascPLOT & ROUND == round)
                sprintf("\n### Round %s \n", round) |> cat()
                insert_image(PATH[idx]) |> cat()
            }
        }else if(.N == 1){
                insert_image(PATH) |> cat()
        }
    }, by = PLOT] 
    
    # avoids unwanted "Empty data.table() message"
    cat()
}
```

